---
title: "R Notebook"
output: html_notebook
---
#Packages and Functions
```{r}
#load Packages
library(dplyr)
library(tidyr)
library(ggplot2)
library(ggrepel)
library(stringr)
library(lubridate)
library(plotly)
library(zoo)
library(purrr)
library(gtools)
library(data.table)
library(randomForest)
#library(rJava)
library(bartMachine)
library(plm)
```

```{r, echo=FALSE, message=FALSE, error=FALSE, warning=FALSE}
#Graphs theme
my_theme <- theme(panel.grid = element_blank(),
        panel.background = element_blank(),
        plot.title = element_text(hjust = 0.5),
        strip.background = element_blank())
```

```{r}
#Functions
##0-1 normalization 
normalize_scalar1 <- function(x, na.rm = T)
                        {x / sqrt(sum(x^2, na.rm = na.rm))}
#not in
'%!in%' <- function(x,y) !('%in%'(x,y))
```

#read, Transform, Aggregate, Merge
##read data
```{r}
#main set 
d <- fread("ondemand_clean.csv") %>% 
  mutate(month_date = as.Date(paste("01-", month, sep = ""), format = "%d-%b-%Y"))

#show data (valid shows only)
show_data <- read.csv("show_data.csv")
show_data$show <- ifelse(show_data$show == "Radiolab Presents: More Perfect",
                           "More Perfect", as.character(show_data$show))
show_data[43,1] <- "Late Night Whenever"
```

##transform data / new variables
```{r}
#keep shows that make sense
d <- d %>% filter(show %in% show_data$show)

#format date variables
d[, grep("date", colnames(d))] <- lapply(d[, grep("date", colnames(d))],
                                        function(x) as.Date(x, format = "%Y-%m-%d"))

#make week date a monday
d$week_date <- floor_date(d$week_date, "weeks") + 1

#extract names of fedd warmers
##by episode, published to podcast
d_eps <- d %>% filter(published_to_podcast == 1) %>% 
  group_by(show, story) %>%
  summarise(story_start_date = story_start_date[1],
            downloads = sum(downloads),
            start_month = format(story_start_date, "%b-%Y"),
            start_year = year(story_start_date),
            start_week = week(story_start_date),
            duration = story_duration[1],
            show_start_date = show_start_date[1]) %>%
  arrange(show, story_start_date) %>% 
  mutate(start_weekday = weekdays(story_start_date))

temp <- d_eps %>%
  filter(show_start_date >= as.Date("2016-03-01"),
         show != "The New Yorker Comment") %>%
  arrange(show, story_start_date) %>%
  mutate(n = row_number()) %>%
  filter(n < 2)

#fix show start date for S01 available shows
d <- d %>% filter(story %!in%  temp$story) %>% group_by(show) %>% 
  mutate(show_start_date = if_else(show_start_date >= min(d$date),
                                  as.Date(min(date)), as.Date(show_start_date)))
#make show start week varible 
d$show_start_week <- floor_date(d$show_start_date, "weeks") + 1
d <- d %>% filter(story %!in%  temp$story) %>% group_by(show, story) %>% 
  mutate(story_start_day_alt = min(date)) 

#complete data for show?
d$show_complete_data <- ifelse(d$show_start_date >= as.Date("2016-03-01"),
                               1,0)
```

##aggregations
```{r}
#complete show reference
complete_shows <- d %>% filter(show_complete_data == 1) %>% 
  group_by(show) %>%
  summarise(start_date = min(show_start_date)) %>%
  arrange(start_date)



##by month, published to podcast
d_month <- d %>% filter(published_to_podcast == 1) %>% 
  group_by(show, month) %>%
  summarise(downloads = sum(downloads),
            n_eps_month = length(which(d_eps$start_month == month[1] & 
                                         d_eps$show == show[1])),
            running_months = difftime(date[1], show_start_date[1], units = "months") + 1,
            avg_ep_length = mean(story_duration),
            show_start_date = show_start_date[1]) %>%
  arrange(show, month_date)

##by week, published to podcast
d_week <- d %>% #filter(published_to_podcast == 1,
                      #story %!in%  temp$story
                      #) %>%
  arrange(show, date) %>%
  #group_by(show) %>% 
 # mutate(show_start_date = if_else(show_start_date >= as.Date("2016-03-01"),
  #                                 min(story_start_day_alt), show_start_date)) %>% 
  group_by(show, week_date) %>%
  summarise(downloads = sum(downloads),
            week_episodes = sum(published_that_day),
            avg_ep_length = mean(story_duration),
            show_start_date = show_start_date[1],
            running_weeks = round(difftime(date[1], show_start_date[1], units = "weeks") + 1, 2)) %>%
  arrange(show, week_date)
  
d_week <- d_week[-(which(d_week$show %in% c("A Piece of Work", "Nancy", "The New Yorker Comment") &
                   d_week$week_episodes == 0 & d_week$running_weeks == 1)), ]
```


##Merge main set with show data
```{r}
d_week <- d_week %>% left_join(show_data)
d_week$show_start_month <- as.factor(lubridate::month(as.Date(d_week$show_start_date), abbr = T))
class(d_week$age_avg)
```



#Models
##case 0: all cases
###Linear
```{r}
d_week$itunes_genre <- relevel(d_week$itunes_genre, "society and culture")
linear0 <- lm(downloads~ week_episodes + avg_ep_length +
                as.factor(show_start_month) +
                running_weeks + station +
                itunes_genre + 
                X12mo_searches_max*running_weeks +
                as.numeric(as.character(age_avg)) +
                klout_avg*running_weeks + twitter_max*running_weeks +
                host_has_previous_show + is_spinoff + is_brand + it_airs, data = d_week)
```

```{r}
summary(linear0)

```


###Random Forest

```{r}
trainingrf0 <- d_week %>% mutate(show_start_month = as.factor(show_start_month)) %>% 
  select(downloads, week_episodes, avg_ep_length, running_weeks, show_start_month,
         running_weeks, station, X12mo_searches_max, age_avg, klout_avg, itunes_genre,
         twitter_max, host_has_previous_show, is_spinoff, is_brand, it_airs)

trf0 <- trainingrf0[complete.cases(trainingrf0),]
trf0$show_start_month <- as.factor(trf0$show_start_month)
rf0 <- randomForest(downloads~ week_episodes + avg_ep_length + show_start_month + running_weeks + station + itunes_genre + 
                X12mo_searches_max*running_weeks + age_avg + klout_avg*running_weeks + twitter_max*running_weeks +
                host_has_previous_show + is_spinoff + is_brand + it_airs, data = trf0, importance = T)
```

```{r, fig.width=10, fig.height= 4}
varImpPlot(rf0)
```

#Fixed effects
```{r}
fixed1 <- plm(downloads~ week_episodes + avg_ep_length +
                as.numeric(show_start_month) +
                running_weeks + station +
                itunes_genre + 
                X12mo_searches_max*running_weeks +
                as.numeric(as.character(age_avg)) +
                klout_avg*running_weeks + twitter_max*running_weeks +
                host_has_previous_show + is_spinoff + is_brand + it_airs, index = c("show", "week_date"), model = "within", data = d_week)
```

```{r}
summary(fixed1)
```



##First Training: incomplete data
```{r}
d_week$age_avg <- as.numeric(as.character(d_week$age_avg))
training1 <- d_week %>% filter(show %!in% complete_shows$show)
trainingrf1 <- d_week %>% filter(show %!in% complete_shows$show) %>%
  select(downloads, week_episodes, avg_ep_length, running_weeks, show_start_month,
         running_weeks, station, X12mo_searches_max, age_avg, klout_avg, itunes_genre,
         twitter_max, host_has_previous_show, is_spinoff, is_brand, it_airs)

trf1 <- trainingrf1[complete.cases(trainingrf1),]
```

###Linear
```{r}
linear1 <- lm(downloads~ week_episodes + avg_ep_length +
                as.numeric(show_start_month) +
                running_weeks + station +
                itunes_genre + 
                X12mo_searches_max*running_weeks +
                as.numeric(as.character(age_avg)) +
                klout_avg*running_weeks + twitter_max*running_weeks +
                host_has_previous_show + is_spinoff + is_brand + it_airs, data = training1)
```

```{r}
summary(linear1)
```


###Random Forest

```{r}
rf1 <- randomForest(downloads~ week_episodes + avg_ep_length + show_start_month + running_weeks + station + itunes_genre + 
                X12mo_searches_max*running_weeks + age_avg + klout_avg*running_weeks + twitter_max*running_weeks +
                host_has_previous_show + is_spinoff + is_brand + it_airs, data = trf1)
```


##First testing: Complete Data
```{r}
testing1 <- d_week %>% filter(show %in% complete_shows$show) %>% group_by(show) %>% filter(running_weeks == min(running_weeks))
testing1_lin <- testing1 %>% filter(itunes_genre %!in% c("comedy", "governent & organizations", "kids & family"))

```

```{r}
a <- data.frame(show = testing1_lin$show, dl = testing1_lin$downloads, pred = predict(linear1, newdata = testing1_lin))
```

```{r}
b <- data.frame(show = testing1$show, dl = testing1$downloads, pred = predict(rf1, newdata = testing1))
```

```{r}
varImpPlot(rf1)
```

###Predict
###MAPE, MSE

##Second Training: Up to "A Piece of Work"
```{r}
training2 <- d_week %>% filter(show %!in% complete_shows[11:16,]$show)
trainingrf2 <- d_week %>% filter(show %!in% complete_shows[11:16,]$show) %>%
  select(downloads, week_episodes, avg_ep_length, running_weeks, show_start_month,
         running_weeks, station, X12mo_searches_max, age_avg, klout_avg, itunes_genre,
         twitter_max, host_has_previous_show, is_spinoff, is_brand, it_airs)

trf2 <- trainingrf2[complete.cases(trainingrf2),]
```

###Linear
```{r}
linear2 <- lm(downloads~ week_episodes + avg_ep_length +
                #show_start_month +
                running_weeks + station +
                #itunes_genre + 
                X12mo_searches_max*running_weeks +
                as.numeric(age_avg) +
                klout_avg*running_weeks + twitter_max*running_weeks +
                host_has_previous_show + is_spinoff + is_brand + it_airs, data = training2)
```

```{r}
summary(linear2)
```


###Random Forest
```{r}
rf2 <- randomForest(downloads~ week_episodes + avg_ep_length + show_start_month + running_weeks + station + itunes_genre + 
                X12mo_searches_max*running_weeks + age_avg + klout_avg*running_weeks + twitter_max*running_weeks +
                host_has_previous_show + is_spinoff + is_brand + it_airs, data = trf2)
```

##Second Testing: After "A Piece of Work"
```{r}
testing2 <- d_week %>% filter(show %in% complete_shows[1:11,]$show) %>% group_by(show) %>% filter(running_weeks == min(running_weeks))
testing2_lin <- testing1 %>% filter(itunes_genre %!in% c("comedy", "governent & organizations", "kids & family"))

```

```{r}
c <- data.frame(show = testing2$show, dl = testing2$downloads, pred = predict(linear2, newdata = testing2))
```

```{r}
d <- data.frame(show = testing2$show, dl = testing2$downloads, pred = predict(rf2, newdata = testing2))
```

```{r}
varImpPlot(rf2)
```

###Predict
###MAPE, MSE