---
title: "Session ID Review"
output:
  html_document:
    df_print: paged
---

```{r, echo=FALSE, message=FALSE, error=FALSE, warning=FALSE}
#load Packages
library(dplyr)
library(tidyr)
library(ggplot2)
library(ggrepel)
library(stringr)
library(lubridate)
```

```{r, echo=FALSE, message=FALSE, error=FALSE, warning=FALSE}
#Graphs theme
my_theme <- theme(panel.grid = element_blank(),
        panel.background = element_blank(),
        plot.title = element_text(hjust = 0.5))

```

```{r, echo=FALSE, message=FALSE, error=FALSE, warning=FALSE}
#read data
d <- read.csv("PlatformEvents.csv")
d <- d %>% arrange(session_id, event_timestamp)
d$event_timestamp <- as.POSIXct(d$event_timestamp)

d2 <- read.csv("platform_events_sessions.csv")
d2$max_timestamp <- as.POSIXct(d2$max_timestamp)
d2$min_timestamp <- as.POSIXct(d2$min_timestamp)
```

```{r, echo=FALSE, message=FALSE, error=FALSE, warning=FALSE}
#New Variables
##last piece of session id
d$sess_id_in_day <- as.numeric(str_extract(d$session_id, "[^_]*$"))
##Action Type
d$action_type <- ifelse(d$action %in% c('set_position', 'start', 'resume',
                                        'skip_15_forward', 'skip_15_back'),
                        'play',
                        ifelse(d$action %in% c('finish', 'pause', 'window_close',
                                               'interrupt'),
                               'stop',
                               ifelse(d$action == 'page_view',
                                      'page_view', NA)))
##day session variable
d$day_session <- str_extract(d$session_id, "^(.*[_])")

d2$sess_id_in_day <- as.numeric(str_extract(d2$session_id, "[^_]*$"))
```

```{r, echo=FALSE, message=FALSE, error=FALSE, warning=FALSE}
#sample
d_small <- sample_n(d, 20000)
```

```{r, echo=FALSE, message=FALSE, error=FALSE, warning=FALSE}
#
session_id_summary <- d %>% group_by(session_id) %>%
  dplyr::summarise(n_parts = max(session_part_id),
            first_action = action[1],
            first_action_type = action_type[1],
            first_action_timestamp = event_timestamp[1],
            last_action = action[length(session_id)],
            last_action_type = action_type[length(session_id)],
            last_action_timestamp = event_timestamp[length(session_id)],
            last_action_in_day = sess_id_in_day[length(session_id)],
            ses_len = difftime(event_timestamp[length(session_id)],
                               event_timestamp[1], units = "mins"),
            is_day_crossover = ifelse(last_action_in_day < 0, 1, 0)
            #,
            #n_user_ids = length(unique(user_id)),
            #n_platf_event_id = length(unique(platform_events_id))
            )
session_id_summary$ses_len <- round(session_id_summary$ses_len, 3)
```

```{r}
#Check odd first/last actions



```

##General

```{r, echo=FALSE, message=FALSE, error=FALSE, warning=FALSE}
ggplot(d %>% group_by(action) %>% 
         count() %>% 
         arrange(desc(n)) %>%
         ungroup() %>% 
         mutate(action = factor(action, levels = action))) +
  labs(title = "Action Frequency") +
  geom_bar(aes(x = action, y = n),
           stat = "identity") +
  my_theme +
  theme(axis.text.x = element_text(angle = 45,
                                   hjust = 1))
```

Is it abnormal that 'set_position' has a higher frequency than 'start'?

```{r, echo=FALSE, message=FALSE, error=FALSE, warning=FALSE}
ggplot(d %>% group_by(action_type) %>% 
         count() %>% 
         arrange(desc(n)) %>%
         ungroup() %>% 
         mutate(action_type = factor(action_type,
                                     levels = action_type))) +
  labs(title = "Type of Action Frequency") +
  geom_bar(aes(x = action_type, y = n),
           stat = "identity") +
  my_theme +
  theme(axis.text.x = element_text(angle = 45,
                                   hjust = 1))
```

There are more 'play' than 'stop' actions in total. I don't really know if we were expecting there to be an equal amount; it would make sense, but there is probably a reason for this. 


```{r, echo=FALSE, message=FALSE, error=FALSE, warning=FALSE}
ggplot(d) +
  labs(title = "Session ID in Day Identifier, Histogram.") +
  geom_histogram(aes(x = sess_id_in_day),
                 binwidth = 0.5,
                 bins = 19) +
  scale_x_continuous(breaks = c(-1:18)) +
  my_theme
```

I think this looks good, the -1 identifiers are a relative minority. Most cases have only 1 or 2 sessions a day. I thought that the crossover identifiers could go to -n if the last session of the day was the nth and then crossed over. For the data, the min value of the identifier is -1.


## Within Session ID

```{r, echo=FALSE, message=FALSE, error=FALSE, warning=FALSE}
ggplot(session_id_summary) +
  labs(title = "Type of First Action. Frequency") +
  geom_bar(aes(x = factor(first_action_type, levels = c("play", "page_view", "stop")))) +
  my_theme +
  theme(axis.text.x = element_text(angle = 45,
                                   hjust = 1))
```

First actions within Session ID are more often "stops" than "plays" 

```{r, echo=FALSE, message=FALSE, error=FALSE, warning=FALSE}
ggplot(session_id_summary %>% 
         group_by(first_action_type,
                  first_action) %>% 
         count() %>% 
         arrange(first_action_type, desc(n)) %>%
         ungroup() %>% 
         mutate(first_action = factor(first_action,
                                      levels = first_action))) +
  labs(title = "First Action. Frequency") +
  geom_bar(aes(x = first_action,
               y = n,
               fill = first_action_type),
           stat = "identity") +
  my_theme +
  theme(axis.text.x = element_text(angle = 45,
                                   hjust = 1))
```

In dissaggregated first actions, "window_close" seems to be the most frequent after "page_view"


```{r, echo=FALSE, message=FALSE, error=FALSE, warning=FALSE}
ggplot(session_id_summary) +
  labs(title = "Type of Last Action. Frequency") +
  geom_bar(aes(x = last_action_type)) +
  my_theme +
  theme(axis.text.x = element_text(angle = 45,
                                   hjust = 1))
```


```{r, echo=FALSE, message=FALSE, error=FALSE, warning=FALSE}
ggplot(session_id_summary %>% 
         group_by(last_action_type,
                  last_action) %>% 
         count() %>% 
         arrange(last_action_type, desc(n)) %>%
         ungroup() %>% 
         mutate(last_action = factor(last_action,
                                      levels = last_action))) +
  labs(title = "Last Action. Frequency") +
  geom_bar(aes(x = last_action,
               y = n,
               fill = last_action_type),
           stat = "identity") +
  my_theme +
  theme(axis.text.x = element_text(angle = 45,
                                   hjust = 1))
```


```{r, echo=FALSE, message=FALSE, error=FALSE, warning=FALSE}
ggplot(session_id_summary %>% filter(ses_len < 50)) +
  labs(title = "Session Length (< 50 mins) Histogram.",
       x = "minutes") +
  geom_histogram(aes(x = ses_len),
                 binwidth = 1) +
  scale_x_continuous(breaks = c(seq(0,50,5))) +
  my_theme

summary(as.numeric(session_id_summary$ses_len))
```

The wide majority of sessions last less than 3 minutes


```{r, echo=FALSE, message=FALSE, error=FALSE, warning=FALSE}
ggplot(session_id_summary %>% filter(ses_len > 50)) +
  labs(title = "Session Length (> 50 mins) Histogram.",
       x = "minutes") +
  geom_histogram(aes(x = ses_len),
                 binwidth = 1) +
  my_theme

```

A minority of cases lasts over 50 minutes, the most extreme session ID lasts about 45 hours

```{r, echo=FALSE, message=FALSE, error=FALSE, warning=FALSE}
ggplot(session_id_summary %>% filter(ses_len <50) %>% 
         group_by(first_action_type,
                  first_action) %>% 
         count() %>% 
         arrange(first_action_type, desc(n)) %>%
         ungroup() %>% 
         mutate(first_action = factor(first_action,
                                      levels = first_action))) +
  labs(title = "Type of First Action (Session Length < 50 mins). Frequency") +
  geom_bar(aes(x = first_action,
               y = n,
               fill = first_action_type),
           stat = "identity") +
  my_theme +
  theme(axis.text.x = element_text(angle = 45,
                                   hjust = 1))
```


```{r, echo=FALSE, message=FALSE, error=FALSE, warning=FALSE}
ggplot(session_id_summary %>% filter(ses_len >50) %>% 
         group_by(first_action_type,
                  first_action) %>% 
         count() %>% 
         arrange(first_action_type, desc(n)) %>%
         ungroup() %>% 
         mutate(first_action = factor(first_action,
                                      levels = first_action))) +
  labs(title = "Type of First Action (Session Length > 50 mins). Frequency") +
  geom_bar(aes(x = first_action,
               y = n,
               fill = first_action_type),
           stat = "identity") +
  my_theme +
  theme(axis.text.x = element_text(angle = 45,
                                   hjust = 1))
```

```{r, echo=FALSE, message=FALSE, error=FALSE, warning=FALSE}
ggplot(session_id_summary %>% filter(ses_len <50) %>% 
         group_by(last_action_type,
                  last_action) %>% 
         count() %>% 
         arrange(last_action_type, desc(n)) %>%
         ungroup() %>% 
         mutate(last_action = factor(last_action,
                                      levels = last_action))) +
  labs(title = "Type of last Action (Session Length < 50 mins). Frequency") +
  geom_bar(aes(x = last_action,
               y = n,
               fill = last_action_type),
           stat = "identity") +
  my_theme +
  theme(axis.text.x = element_text(angle = 45,
                                   hjust = 1))
```

```{r, echo=FALSE, message=FALSE, error=FALSE, warning=FALSE}
ggplot(session_id_summary %>% filter(ses_len >50) %>% 
         group_by(last_action_type,
                  last_action) %>% 
         count() %>% 
         arrange(last_action_type, desc(n)) %>%
         ungroup() %>% 
         mutate(last_action = factor(last_action,
                                      levels = last_action))) +
  labs(title = "Type of last Action (Session Length > 50 mins). Frequency") +
  geom_bar(aes(x = last_action,
               y = n,
               fill = last_action_type),
           stat = "identity") +
  my_theme +
  theme(axis.text.x = element_text(angle = 45,
                                   hjust = 1))
```


##Crossover vs not 

```{r, echo=FALSE, message=FALSE, error=FALSE, warning=FALSE}
ggplot(session_id_summary %>% 
         mutate(is_day_crossover = ifelse(is_day_crossover == 1, "crossover", "not_crossover"))) +
  labs(title = "Type of First Action. By Crossover. Frequency") +
  geom_bar(aes(x = first_action_type)) +
  facet_wrap(~is_day_crossover, scales = "free_y") +
  my_theme +
  theme(axis.text.x = element_text(angle = 45,
                                   hjust = 1))
```


```{r, echo=FALSE, message=FALSE, error=FALSE, warning=FALSE}
ggplot(session_id_summary %>% 
         mutate(is_day_crossover = ifelse(is_day_crossover == 1, "crossover", "not_crossover"))) +
  labs(title = "Type of Last Action. By Crossover. Frequency") +
  geom_bar(aes(x = last_action_type)) +
  facet_wrap(~is_day_crossover, scales = "free_y") +
  my_theme +
  theme(axis.text.x = element_text(angle = 45,
                                   hjust = 1))
```


#Above 1 Minute Duration Sessions 


```{r, echo=FALSE, message=FALSE, error=FALSE, warning=FALSE}
ggplot(session_id_summary %>% filter(ses_len > 1)) +
  labs(title = "Type of First Action. Frequency") +
  geom_bar(aes(x = factor(first_action_type, levels = c("play", "page_view", "stop")))) +
  my_theme +
  theme(axis.text.x = element_text(angle = 45,
                                   hjust = 1))
```



```{r, echo=FALSE, message=FALSE, error=FALSE, warning=FALSE}
ggplot(session_id_summary %>% filter(ses_len > 1) %>% 
         group_by(first_action_type,
                  first_action) %>% 
         count() %>% 
         arrange(first_action_type, desc(n)) %>%
         ungroup() %>% 
         mutate(first_action = factor(first_action,
                                      levels = first_action))) +
  labs(title = "First Action. Frequency") +
  geom_bar(aes(x = first_action,
               y = n,
               fill = first_action_type),
           stat = "identity") +
  my_theme +
  theme(axis.text.x = element_text(angle = 45,
                                   hjust = 1))
```


```{r, echo=FALSE, message=FALSE, error=FALSE, warning=FALSE}
ggplot(session_id_summary %>% filter(ses_len > 1)) +
  labs(title = "Type of Last Action. Frequency") +
  geom_bar(aes(x = factor(last_action_type, levels = c("play", "page_view", "stop")))) +
  my_theme +
  theme(axis.text.x = element_text(angle = 45,
                                   hjust = 1))
```


```{r, echo=FALSE, message=FALSE, error=FALSE, warning=FALSE}
ggplot(session_id_summary %>%  filter(ses_len > 1) %>% 
         group_by(last_action_type,
                  last_action) %>% 
         count() %>% 
         arrange(last_action_type, desc(n)) %>%
         ungroup() %>% 
         mutate(last_action = factor(last_action,
                                      levels = last_action))) +
  labs(title = "Last Action. Frequency") +
  geom_bar(aes(x = last_action,
               y = n,
               fill = last_action_type),
           stat = "identity") +
  my_theme +
  theme(axis.text.x = element_text(angle = 45,
                                   hjust = 1))
```

###Crossover vs not 

```{r, echo=FALSE, message=FALSE, error=FALSE, warning=FALSE}
ggplot(session_id_summary %>%  filter(ses_len > 1) %>% 
         mutate(is_day_crossover = ifelse(is_day_crossover == 1, "crossover", "not_crossover"))) +
  labs(title = "Type of First Action. By Crossover. Frequency") +
  geom_bar(aes(x = first_action_type)) +
  facet_wrap(~is_day_crossover, scales = "free_y") +
  my_theme +
  theme(axis.text.x = element_text(angle = 45,
                                   hjust = 1))
```


```{r, echo=FALSE, message=FALSE, error=FALSE, warning=FALSE}
ggplot(session_id_summary%>%  filter(ses_len > 1) %>% 
         mutate(is_day_crossover = ifelse(is_day_crossover == 1, "crossover", "not_crossover"))) +
  labs(title = "Type of Last Action. By Crossover. Frequency") +
  geom_bar(aes(x = last_action_type)) +
  facet_wrap(~is_day_crossover, scales = "free_y") +
  my_theme +
  theme(axis.text.x = element_text(angle = 45,
                                   hjust = 1))
```

```{r}
summary(as.numeric((session_id_summary %>% filter(is_day_crossover == 1))$ses_len))
```

#Aggregate Data

```{r}
#make duration variable 
d2$duration <- difftime(d2$min_timestamp, d2$max_timestamp, units = "mins")

ggplot(d2 %>% filter(duration < 100))+
  geom_histogram(aes(x = duration),
                     binwith = 1,
                     bins = 300) +
  my_theme
```

```{r}
summary(as.numeric(d2$duration))
summary(as.numeric(session_id_summary$ses_len))
```



#Checking 0 Length cases 

```{r}
nrow(session_id_summary %>% filter(ses_len != 0)) /
  nrow(session_id_summary %>% filter(ses_len == 0))

```

```{r}
nrow(d2 %>% filter(duration != 0)) /
  nrow(d2 %>% filter(duration == 0))

```










